### 2016-05-05 ch03-lrm/truncation.tex
Some cleanup here and there
Some cleanup in truncation

\label{sec:nonparametric:truncation}
\researchBasedOn[This section]{Lumori2014TIM}
%TODO: {labels into standardized form}

% This section uses nonparametric models in conjunction with the \gls{LPM}~\citep{Pintelon2010LPM1} to estimate the complex-valued \gls{FRF} in a linear \gls{LS} fashion.
% The advantage of using the nonparametric models is basically that the `models` often are very simple and hence require little user interaction to construct.
% Among the current nonparametric methods, the \gls{LPM} method is chosen due to its superiority over the classical nonparametric methods (windowing) in suppressing leakage errors~\citep{Bendat1993,Oppenheim1983,Schoukens2006LPM}. 
% Consequently, the \gls{LPM} estimate of the \gls{FRF} is not hampered by leakage errors \citep{Pintelon2010LPM1,Schoukens2009LPM}. 
% Note that a more general discussion on the \gls{LPM}, applied to noisy, real-valued data is found in \citep{Fan1996} with a specific focus on parametric tuning.

@@

% \subsection*{Theory on the LPM and Windowing}
% \label{sec:nparam:trunc:theoryLPMandWindowing}

% Details of the theory pertaining to this section may be found in \citet{Schoukens2009LPM}, basic aspects of which are presented below.

% \subsubsection{Spectral Nonparametric Model and Assumptions}

% Using convolution, the noiseless output signal in \figref{fig:nparam:blockH0} is $y_0(t) = g_0(t)*u_0(t)$, where the signals are sampled at $t = 0, 1, 2,...,N-1$ and the excitation signal $u_0(t)$ is random noise. 
% As shown in \citet{Pintelon2012} and \citet{Pintelon1997ARB} the \gls{DFT} spectra of the sampled signals at the frequency $\omega_k$ are related as follows:
% \begin{equation}\label{eq:nparam:trunc:LPM:leakage}
% Y_0(k)=G_0(\omega_k)U_0(k)+T_G(\omega_k)
% \end{equation}
% where $U_0(k)$ is the \gls{DFT} spectrum of the random noise excitation,  $G_0(\omega_k)$ is the transfer function and $T_G(\omega_k)$ is the transients leakage term.
% In this section, the \gls{DFT} of a signal $x(t)$ is defined explicitly as~\citep{Oppenheim1983}

% \begin{equation}\label{eq:defDFT}
% X(\omega_k) = \frac{1}{\sqrt{N}}\sum_{t=0}^{N-1}x(t)e^{-\frac{j2\pi kt}{N}}
% \end{equation}

% The following assumptions are consistent with \citep{Schoukens2009LPM} for an \gls{OE} setup (\figref{fig:nparam:blockH0}):
% \begin{assumption}
% The \gls{SISO} system $G_0$ and  the transients leakage spectrum $T_G$ are both smooth functions of the frequency.
% \end{assumption}

@@

% The smooth characteristics of $G_0(\omega_k)$ and $T(\omega_k)$ allow for the use of the Taylor series representations at the frequency lines $k+r$ in a frequency domain window of width $2n+1$, with a remainder $\bigO{\left(\frac{r}{N}\right)^{R+1}}$ of the first order Taylor series expansion \citep{Schoukens2009LPM}, \latin{viz}:
% \begin{align}\label{lpmGTaylorS}
% G_0(\omega_{k+r})&=G_0(\omega_k)+\sum_{p=1}^{R}g_p(k)r^p+
% \bigO{\left(\frac{r}{N}\right)^{R+1}},
% \\
% \label{lpmTTaylorS}
% T(\omega_{k+r})&=T(\omega_k)+\sum_{p=1}^{R}t_p(k)r^p+N^{-1/2}
% \bigO{  \left(\frac{r}{N}\right)^{R+1}  }
% ,
% \\
% \text{where}&\ r\in\mathbb{W}_n,\quad \mathbb{W}_n = \{-n,-n+1,\dots,n\}
% \end{align}

% The choice of the tunable parameters $n$ and $R$ is discussed later on.

% Neglecting the remainder, the Taylor series approximation of \eqref{lpm1spectra} can be written in a matrix vector form, \latin{viz}:
% \begin{equation}\label{lpmGTMatrix}
% Y(k+r)\approx{K(k, r)\theta(k)+V_0(k+r)}
% \end{equation}
% where $\theta(k)\in\mathbb{C}^{2R+2}$ is the column vector of the  parameters ($G_0(\omega_k)$, $T(\omega_k)$, $g_p$, $t_p$, and $p = 1, 2, ...., R$) and $K(k, r)$ is the row vector of their respective coefficients:
% \begin{align}
% K(k,r) = &\left[
% \begin{matrix}
% U_0(k+r) & U_0(k+r)r & \dots
% \end{matrix}
% \right.
% \\
% &\quad\left.\begin{matrix}U_0(k+r)r^R & 1& r&\dots&r^R\end{matrix}\right]\nonumber
% \end{align}

% Equation \eqref{lpmGTMatrix} can be written in the more compact (stacked vectors) form:
% \begin{equation}\label{lpmGTMatrixStack}
% Y_n(k)\approx K_n(k)\theta(k)+V_n(k)
% \end{equation}
% where $Y_n(k), V_n(k)\in\mathbb{C}^{2n+1}$, and $K_n(k)\in\mathbb{C}^{(2n+1)\times(2R+2)}$ are the stacked values of $Y(k+r)$, $K(k,r)$, and $V_0(k+r)$, respectively, for $r\in\mathbb{W}_n$.
% The estimates of the parameters $\theta(k)$ are obtained as the minimizers of the least squares problem, \latin{viz}:
% \begin{align}\label{eq:LSsolutionLPM}
% \hat\theta(k) = \argmin_{\theta(k)} \left\|Y_n(k) - K_n(k)\theta(k)\right\|_2^2
% \end{align}

% The formulation of equation \eqref{eq:LSsolutionLPM} must take into account the fact that  a full rank of $K_n$ requires $n \geq R + 1$. The choice $n = R + 1$ results in the highest variance on the estimate, but with the smallest interpolation error \citep{Schoukens2009LPM}.

% The values $R = 2$, and $n=R+1=3$ will be used in the \gls{LPM}.
In this section, the $\lpm{3,2,2}$ is used.
However, since the techniques discussed in this section are not specifically geared towards the \gls{LPM}, one can just as well use the concepts with other non-parametric \gls{FRF} estimators such as the \gls{LRM}.

% When substituted into equations \eqref{lpmGTaylorS} and \eqref{lpmGTMatrix},  these yield the following quadratic local polynomials of $G_0$ and $T$ around a central frequency $\omega_{k}$:
% \begin{align}
% G_0(\omega_{k+r})
%   &\approx 
%   G_0(\omega_k)+g_1r+g_2r^2
%   \label{eq:lpm:expansionG:quadratic}
% \\
%   T(\omega_{k+r})
%     &\approx 
%     T(\omega_k)+t_1r+t_2r^2
%     \label{eq:lpm:expansionT:quadratic}
% \end{align}
% with ($r\in\mathbb{W}_3$) and the vector of parameters:
% \begin{equation}\label{lpmThetaEst}
% \theta^{\TT}(k)=\left[G_0(\omega_k) \ g_1 \ g_2\  T_G(k)\  t_1 \ t_2\right]
% \end{equation}
% The \gls{LPM} estimate of the \gls{FRF} is the first element of the estimate $\hat\theta(k)$, obtained from \eqref{eq:LSsolutionLPM}, \latin{viz}:
% \begin{equation}
% \hat{G}_{\LPM}(\omega_k) = \hat\theta_1(k)
% \end{equation}

@@

The estimated impulse response function  $\hat g_{{\LPM}\setminus \mathrm{DC}}(t)$ is obtained explicitly via the \gls{IDFT} of the \gls{LPM}-estimate $\hat G_{\LPM}(\omega_k)$ of the \gls{FRF}, \latin{viz}:
\begin{align}\label{eq:impRespiDFT}
\hat g_{{\LPM}\setminus \mathrm{DC}}(t) 
&= 
\frac{1}{\sqrt{N}}
\sum_{k=1}^{N-1}
\hat G_{\LPM}(\omega_k)e^{\frac{j2\pi kt}{N}}
\end{align}
where the estimated \gls{FRF} in the frequency band between  the Nyquist and the sample frequencies is obtained as follows:
\begin{align}
\hat G_{\LPM}(\omega_{N-k}) = \overline{\hat G_{\LPM}(\omega_k)},\ \text{for}\ k=1,\dots,N/2
\end{align}
(with $\conj{\hat G_{\LPM}}$ the complex conjugate of $\hat G_{\LPM}$) to ensure $\hat g_{\LPM}(t)$ to be real.

Smoothing of the estimated \gls{FRF} requires the correct estimate of the impulse response. 
Unfortunately, the \gls{LPM} presented in \secref{sec:nparam:LRM} and in \citep{Schoukens2009LPM} does not estimate the \gls{DC} value of the \gls{FRF}, hence the subscript ``$\setminus\mathrm{DC}$'' in equation \eqref{eq:impRespiDFT}. 
Consequently, the mean value of the corresponding estimated impulse response given in equation \eqref{eq:impRespiDFT} is not correct. 
This limitation is lifted by developing a simple estimator of the mean value, presented below.

@@

\begin{enumerate}
\item Compute the mean value $m_{g10}$ of the last 10\% of the impulse response $\hat g_{{\LPM}\setminus \mathrm{DC
}}(t)$, estimated by the \gls{LPM}, as %$\hat g_{\LPM}(t)$ as
\begin{align}
   m_{g10} 
   = 
   \frac{1}{\lceil0.1N - 1\rceil}
   \sum_{t=\lfloor0.9N\rfloor}^{N-1}
      \hat g_{{\LPM}\setminus \mathrm{DC}}(t)
\end{align}

\item Next, subtract the computed mean value $m_{g10}$ from $\hat g_{{\LPM}\setminus \mathrm{DC
}}(t)$ to obtain the improved impulse response $\hat g_{\LPM}(t)$, \latin{viz}:
\begin{align}
\hat g_{\LPM}(t) = \hat g_{{\LPM}\setminus \mathrm{DC}}(t) - m_{g10},\ \text{for}\ t=0,1,\dots,N-1
\end{align}
\end{enumerate}

\subsection{Determining the Truncation Time}
In this section, two methods will be discussed to determine the truncation time, i.e. the time beyond which the measured impulse response cannot be separated from the noise level.
Specifically, the first method uses an $F$-test to distinguish between signal and noise.
The second method fits an exponential through a well-chosen subset of the data to obtain $\truncTime$.

@@

\subsubsection{Truncation of the Impulse Response Via Exponential Fitting}
\label{sec:nonparametric:truncation:exponentialfit}

>>>>

\label{sec:nonparametric:truncation}
\researchBasedOn[This section]{Lumori2014TIM}
%FIXME: {labels into standardized form}

@@

@@

In this section, the $\lpm{3,2,2}$ is used.
However, since the techniques discussed in this section are not specifically geared towards the \gls{LPM}, one can just as well use the concepts with other non-parametric \gls{FRF} estimators such as the \gls{LRM}.

@@

The estimated impulse response function  $\hat g_{{\LPM}\without \DC}(t)$ is obtained explicitly via the \gls{IDFT} of the \gls{LPM}-estimate $\hat G_{\LPM}(\omega_k)$ of the \gls{FRF}, \latin{viz}:
\begin{align}\label{eq:impRespiDFT}
\hat g_{{\LPM}\without \DC}(t) 
&= 
\frac{1}{\sqrt{N}}
\sum_{k=1}^{N-1}
\hat G_{\LPM}(\omega_k)e^{\frac{j2\pi kt}{N}}
\end{align}
where the estimated \gls{FRF} in the frequency band between  the Nyquist and the sample frequencies is obtained as follows:
\begin{align}
\hat G_{\LPM}(\omega_{N-k}) = \overline{\hat G_{\LPM}(\omega_k)},\ \text{for}\ k=1,\dots,N/2
\end{align}
(with $\conj{\hat G_{\LPM}}$ the complex conjugate of $\hat G_{\LPM}$) to ensure $\hat g_{\LPM}(t)$ to be real.

Smoothing of the estimated \gls{FRF} requires the correct estimate of the impulse response. 
Unfortunately, the \gls{LPM} presented in \secref{sec:nparam:LRM} and in \citep{Schoukens2009LPM} does not estimate the \gls{DC} value of the \gls{FRF}, hence the subscript ``$\without\DC$'' in equation \eqref{eq:impRespiDFT}. 
Consequently, the mean value of the corresponding estimated impulse response given in equation \eqref{eq:impRespiDFT} is not correct. 
This limitation is lifted by developing a simple estimator of the mean value, presented below.

@@

\begin{enumerate}
\item Compute the mean value $m_{g10}$ of the last 10\% of the impulse response $\hat g_{{\LPM}\without \DC}(t)$, estimated by the \gls{LPM}, as %$\hat g_{\LPM}(t)$ as
\begin{align}
   m_{g10} 
   = 
   \frac{1}{\lceil0.1N - 1\rceil}
   \sum_{t=\lfloor0.9N\rfloor}^{N-1}
      \hat g_{{\LPM}\without \DC}(t)
\end{align}

\item Next, subtract the computed mean value $m_{g10}$ from $\hat g_{{\LPM}\without \DC}(t)$ to obtain the improved impulse response $\hat g_{\LPM}(t)$, \latin{viz}:
\begin{align}
\hat g_{\LPM}(t) = \hat g_{{\LPM}\without \DC}(t) - m_{g10},\ \text{for}\ t=0,1,\dots,N-1
\end{align}
\end{enumerate}

\subsection{Determining the Truncation Time of the Impulse Response}
In this section, two methods will be discussed to determine the truncation time, i.e. the time beyond which the measured impulse response cannot be separated from the noise level.
Specifically, the first method uses an $F$-test to distinguish between signal and noise.
The second method fits an exponential through a well-chosen subset of the data to obtain $\truncTime$.

@@

\subsubsection{Truncation Via Exponential Fitting}
\label{sec:nonparametric:truncation:exponentialfit}
